---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2026.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2026.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2026.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

ANSWER HERE
The breadth of the yearly salary data is quite large — ignoring the zeroes (for now), salaries range from $59,661/year to $197,200/year between employees and detailees. Some are part-time, but still receive significant salaries. A snafu for analysts are the employees who report $0 in salaries — what on earth is going on here? Marco Rubio and Steve Witkoff are two such employees, telling me that their salaries are far greater but under a different umbrella of payment through their greater positions. In any case, these zeroes present as outliers in the data, skewing our results.

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
wh_salaries |> summarize(mean_salary = mean(SALARY), median_salary = median(SALARY))
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries |> summarize(mean_salary = mean(SALARY), median_salary = median(SALARY), min_salary = min(SALARY), max_salary = max(SALARY))
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

ANSWER HERE
The data in total vary from $0 to nearly $200,000 — a large margin. Looking at the IQR of the data & the upper & lower bounds (-$5,822.50 to $217,093.50) no values are technically outliers but the $0 salaries are quite questionable & raise many inquiries — how much are these guys actually getting paid?

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

ANSWER HERE
I think the best story here is how some members of the White House are reporting zero dollars in salary. As a journalist, my first step would be to investigate where the real value of their pay actually comes from: does their income spring from a different role they play in the federal government? Where is this reported? Are they not reporting income because they're high-ranking figures? Do they receive money in other ways? Do they donate it to charity?
