---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities |>
  mutate(violent_rate_2019 = violent2019 / pop2019 * 10000) |>
  mutate(property_rate_2019 = property2019 / pop2019 * 10000) |>
  mutate(violent_rate_2020 = violent2020 / pop2020 * 10000) |>
  mutate(property_rate_2020 = property2020 / pop2020 * 10000)
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

ANSWER HERE
In 2019 and 2020, Baltimore and Ocean City were the two cities with the highest violent crime rates per 10,000 residents, respectively, and Ocean City was the city with the highest property crime rate per 10,000 residents for both years. Burkittsville, Sharpstown, Woodsboro, Barclay, Barnesville, Betterton, Eagle Harbor, Lonaconing, Marydel, Midland, New Market, North Brentwood, Sudlersville, Trappe, Vienna, Westernport... a number of cities in both 2019 and 2020 experienced no violent crimes and no property crimes. One trend I see is that crime slipped significantly in many cities between 2019 and 2020 — here's a possible story, but one that's already been written before. By county, I do not see any obvious places that host more violent crimes, but once one scrolls down the list, Prince George's County and Washington County appear more frequently than others. It also appears that cities with large populations are more likely to report a large crime rate, but this is not a rule and there are several notable exceptions — Bowie and violent crime, for example.

### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |>
  mutate(violent_pct_change_rates = violent_rate_2020 - violent_rate_2019 / violent_rate_2020 * 100) |>
  mutate(property_pct_change_rates = property_rate_2020 - property_rate_2019 / property_rate_2020 * 100)
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

ANSWER HERE
Most cities seem to have experienced dips in both property crimes and violent crimes from 2019 to 2020, but according to the data, the city of Frederick experienced a 2,080% decrease in violent crimes and a 985% decrease in property crimes. How could this possibly happen? What factor(s) would cause crime to fall by such a huge margin in two different areas of criminality? I think writing up this outlier is more interesting than reporting on the general trend of falling crime rates because in 2026, most people (I should hope most all people) know what the COVID-19 pandemic did as a general rule to instances of crime, but these Frederick figures are totally out of left field. I think one could also report on disparities of change between violent crimes and property crimes: for example, Elkton saw a 6% decrease in violent crimes but a 358% increase in property crimes from 2019 to 2020. Why?

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

ANSWER HERE
Despite broad decreases in violent and property crime across many Maryland cities from 2019 to 2020, the city of Frederick saw an astonishing 2,080% decrease in violent crimes and a 985% decrease in property crimes across the period, according to state data.

In 2019, the city of Frederick reported 309 violent crimes, while in 2020, it reported only 15. Simultaneously, in 2019, the city reported 1,401 property crimes, while in 2020, it reported only 141. Other Maryland cities also saw decreases in both measured varieties of crime from 2019 to 2020, such as Brunswick, which reported a 94% decrease in violent crimes and a 123% decrease in property crimes. Yet no other city reported such massive percentage changes figures as Frederick did.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
